---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv")
head(activity)
```

## What is mean total number of steps taken per day?
```{r}
library(dplyr)
head(table(activity$date))
bydate <- group_by(activity, date)
step_per_day <- summarize(bydate, steps=sum(steps))
mean(step_per_day$steps, na.rm=TRUE)
median(step_per_day$steps, na.rm=TRUE)
```
  
The mean of the total steps taken per day is 10,766.19.  
The median of the total steps taken per day is 10,765.

## What is the average daily activity pattern?
```{r}
head(table(activity$interval))
byinterval <- group_by(activity, interval)
step_by_interval <- summarize(byinterval, steps=mean(steps, na.rm=TRUE))
plot(step_by_interval, type="l", ylab="Average Steps in Interval",
     main="Average Daily Activity Pattern")
step_by_interval[step_by_interval$steps == max(step_by_interval$steps), ]
step_by_interval[65:80,]
```
  
- activities start between 5am and 6am.  
- the interval with the max average steps is 8:35am.

## Imputing missing values
Where steps is missing, the average steps for that particular interval (calcuated above)
were used.
```{r}
table(is.na(activity$steps))
#step_by_interval$interval <- as.factor(step_by_interval$interval)
#activity$interval <- as.factor(activity$interval)
imactivity <- merge(activity, step_by_interval, by.x="interval", by.y="interval" )
imactivity$steps.x[is.na(imactivity$steps.x)] <- 
    imactivity$steps.y[is.na(imactivity$steps.x)]
bydate <- group_by(imactivity, date)
step_per_day <- summarize(bydate, steps=sum(steps.x))    
hist(step_per_day$steps, breaks=20, main="Histogram of Steps per Day", 
     xlab = "Steps per Day")
mean(step_per_day$steps)
median(step_per_day$steps)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
imactivity$weekdays <- as.factor(weekdays(as.Date(imactivity$date)))
imactivity$weekdays <- factor(imactivity$weekdays, levels(imactivity$weekdays)[c(2,6,7,5,1,3,4)])
table(imactivity$weekdays)

byweekdays <- group_by(imactivity, weekdays, interval)
step_by_iw <- summarize(byweekdays, steps=mean(steps.x))
library(ggplot2)
qplot(interval, steps, data=step_by_iw, facets=weekdays~., geom=c("line"))
```
  
- There are not significant different between weekdays and weekends.    
- It appears that walking activities start later after Thursday.  
- Monday appears to be the most active day of the week.  
- Sunday appears to be the less active day of the week.  